Solving the maximum independant set problem by generic programming einsum networks \thanks{\funding{...}}

### 2021-07-29 paper/paper.tex
switch to SIAM

%\documentclass[aps,twocolumn,longbibliography,english,superscriptaddress]{revtex4-1}
\documentclass{article}
%\usepackage{iclr2021_conference}
\usepackage{natbib}
%\documentclass[a4paper,superscriptaddress,11pt]{article}
\pdfoutput=1
\usepackage[colorlinks=true,urlcolor=blue,citecolor=blue,linkcolor=blue]{hyperref}
\usepackage[english]{babel}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{amssymb}
\usepackage{tabularx}
\usepackage{quoting}
\usepackage{upquote}
\usepackage{subcaption}
\usepackage{multicol}
\usepackage[framemethod=TikZ]{mdframed}
\usetikzlibrary{shapes}
\usepackage{wrapfig}
%\usepackage{caption}
%\usepackage[plain]{algorithm}
\usepackage[linesnumbered, ruled, vlined]{algorithm2e}
\usepackage{algpseudocode}
\usepackage{rotating}
%\usepackage{cite}
\usepackage{booktabs}
%\usepackage{unicode-math}
%\usepackage{algorithm}% http://ctan.org/pkg/algorithm
%\usepackage{algpseudocode}% http://ctan.org/pkg/algpseudocode
\usepackage{xcolor}% http://ctan.org/pkg/xcolor
\makeatletter
\newsavebox{\@brx}
\newcommand{\llangle}[1][]{\savebox{\@brx}{\(\m@th{#1\langle}\)}%
  \mathopen{\copy\@brx\kern-0.5\wd\@brx\usebox{\@brx}}}
\newcommand{\rrangle}[1][]{\savebox{\@brx}{\(\m@th{#1\rangle}\)}%
  \mathclose{\copy\@brx\kern-0.5\wd\@brx\usebox{\@brx}}}
\makeatother
\usepackage{bbm}
\usepackage{jlcode}
\usepackage{graphicx}
\usepackage{amsmath,color,amsthm}
\usepackage{mathrsfs}
\usepackage{float}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{indentfirst}
\usepackage{txfonts}
\usepackage[epsilon, tsrm, altpo]{backnaur}

@@

%%%%%% Shortcut related
\newcommand{\<}{\langle}
\renewcommand{\>}{\rangle}
\newcommand{\out}{{\vx^L}}
\newcommand{\inp}{{\vx^0}}
\newcommand{\cquad}{{{ }_{\quad}}}
\newcommand{\pluseq}{\mathrel{+}=}
\newcommand{\minuseq}{\mathrel{-}=}
\newcommand{\vx}{{\mathbf{x}}}
\newcommand{\vg}{{\mathbf{g}}}
\newcommand{\vp}{{\mathbf{p}}}
\newcommand{\vy}{{\mathbf{y}}}
\newcommand{\Var}{{\mathrm{Var}}}
\newcommand{\Mean}{{\mathrm{E}}}
\newcommand{\vvalue}{{\texttt{value}}}
\newcommand{\grad}{{\texttt{grad}}}
\newcommand{\parameter}{{\texttt{parameter}}}
%%%%%% Convention related
\newcommand{\SWAP}{{\rm SWAP}}
\newcommand{\CNOT}{{\rm CNOT}}
\newcommand{\bigO}{{\mathcal{O}}}
\newcommand{\X}{{\rm X}}
\renewcommand{\H}{{\rm H}}
\newcommand{\Rx}{{\rm Rx}}
\renewcommand{\v}[1]{{\bf #1}}
\newcommand{\dataset}{{\mathcal{D}}}
\newcommand{\wfunc}{{\psi}}
\newcommand{\SU}{{\rm SU}}
\newcommand{\UU}{{\rm U}}
\newcommand{\thetav}{{\boldsymbol{\theta}}}
\newcommand{\gammav}{{\boldsymbol{\gamma}}}
\newcommand{\thetai}{{\theta^\alpha_l}}
\newcommand{\Expect}{{\mathbb{E}}}
\newcommand{\Tr}{{\rm Tr}}
\renewcommand{\cite}[1]{{\citep{#1}}}
\newcommand{\etc}{{\it etc~}}
\newcommand{\etal}{{\it etal~}}
\newcommand{\xset}{\mathbf{X}}
\newcommand{\fl}{\texttt{fl}}
\newcommand{\pdata}{\mathbf{\pi}}
\newcommand{\q}{\mathbf{q}}
\newcommand{\epdata}{\mathbf{\hat{\pi}}}
\newcommand{\gammaset}{\boldsymbol{\Gamma}}
\newcommand{\ei}{{\mathbf{e}_l^\alpha}}
\newcommand{\vtheta}{{\boldsymbol{\theta}}}
\newcommand{\sigmag}{{\nu}}
\newcommand{\sigmai}[2]{{\sigma^{#2}_{#1}}}
\newcommand{\qi}[1]{{q^{\alpha_{#1}}_{#1}}}
\newcommand{\BAS}{Bars-and-Stripes}
\newcommand{\circled}[1]{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {#1}}}}
\newcommand{\qexpect}[1]{{\left\langle #1\right\rangle}}
\newcommand{\expect}[2]{{\mathop{\mathbb{E}}\limits_{\substack{#2}}\left[#1\right]}}
\newcommand{\var}[2]{{\mathop{\mathrm{Var}}\limits_{\substack{#2}}\left(#1\right)}}
\newcommand{\pshift}[1]{{p_{\thetav+#1}}}
\newcommand{\upcite}[1]{\textsuperscript{\cite{#1}}}
\newcommand{\Eq}[1]{Eq.~(\ref{#1})}
\newcommand{\Fig}[1]{Fig.~\ref{#1}}
\newcommand{\Lst}[1]{Listing.~\ref{#1}}
\newcommand{\Tbl}[1]{Table~\ref{#1}}
\newcommand{\Sec}[1]{Sec.~\ref{#1}}
\newcommand{\App}[1]{Appendix~\ref{#1}}
\newcommand{\bra}[1]{\mbox{$\left\langle #1 \right|$}}
\newcommand{\ket}[1]{\mbox{$\left| #1 \right\rangle$}}
\newcommand{\braket}[2]{\mbox{$\left\langle #1 | #2 \right\rangle$}}
\newcommand{\tr}[1]{\mathrm{tr}\mbox{$\left[ #1\right]$}}

@@

\newtheorem{theorem}{\textit{Theorem}}
\newtheorem{corollary}{\textit Branching Rule}
\theoremstyle{definition}\newtheorem{definition}{\textit{Definition}}

@@

%\iclrfinalcopy
\begin{document}
\title{Solving the maximum independant set problem by generic programming einsum networks}

%\author{Jin-Guo Liu\\
%Harvard University\\
%\texttt{jinguoliu@g.harvard.edu}\\
%\AND
%Xun Gao\\
%Harvard University\\
%\texttt{xungao@g.harvard.edu}\\
%}
\maketitle

\begin{abstract}
	Solving the maximum independent set size problem by mapping the graph to an einsum network. 
    By contracting the einsum network with generic element types, we show how to obtain the maximum independent set size,
    the independence polynomial and optimal configurations.
\end{abstract}

\section{Introduction}
MIS problem is hard~\cite{Hastad1996}.
Branching algorithms can solve the MIS problem in $1.1893^n n ^{O(1)}$ time ~\cite{Xiao2017}.
Previous dynamic programming approach~\cite{Fomin2013} can reduce the complexity of computing to $2^{tw(G)}$.
A set is independent if and only if it is a clique in the graph's complement, so the two concepts are complementary.
A set is independent if and only if its complement is a vertex cover.[4] Therefore, the sum of the size of the largest independent set $\alpha (G)$ and the size of a minimum vertex cover $\beta (G)$ is equal to the number of vertices in the graph.

\section{A short introduction to einsum networks}

The word ``einsum'' is a shorthand for Einstein's summation, however, modern einsum notation in program is actually invented by a group of programmers.
Einstein's notation is originally proposed as a generalization to of multiplication between two matrices to the contraction between multiple tensors.
Let $A, B$ be two matrices, the matrix multiplication is defined as $C_{ik} = \sum_{j}A_{ij}B_{jk}$.
It is denoted as $C_i^k = A_i^j B_j^k$ in the Einstein's original notation, where the paired subscript and superscript $j$ is a dummy index summed over.
An example of tensor networks is $C_i^l = A_{ij}^k B^l_k V^j$.
One can map a tensor network to a muti-graph with open edges by viewing a tensor in the expression on the right hand side as a vertex in a graph, a label pairing two tensors as an edge, and the remaining labels as open edges.
We get the graphical notation as the following.

@@

\begin{algorithm}[!ht]
    \small
    \SetAlgoNoLine
    \LinesNumbered
    Let $P = 1$, vector $X = (0,1,2,\ldots,m)$, matrix $\hat X_{ij} = X_i^j$, where $i,j = 0, 1, \ldots m$\;
    \While{true}{
        compute the largest prime $p$ that $\gcd(p, P) = 1 \land p \leq 2^W$\;
        compute the tensor network contraction on $GF(p)$ and obtain $Y = (y_0, y_1, \ldots , y_m) \pmod p $\;
        $A_p = (a_0, a_1, \ldots, a_m) \pmod p = {\rm gaussian\_elimination}(\hat X, Y \pmod p) $\;
        $A_{P\times p} = {\rm chinese\_remainder}(A_P, A_p)$\;
        \If{$A_P = A_{P \times p}$}{
            \Return $A_P$ \tcp*[l]{converged}
        }
        $P = P \times p$\;
    }\caption{Compute independence polynomial exactly without integer overflow}\label{alg:finitefield}
\end{algorithm}

@@

\begin{theorem}
    A MIS-compact tropical tensor is optimal, i.e. none of its none zero entries can be removed without accessing global information.
\end{theorem}

@@

\bibliographystyle{iclr2021_conference}
\bibliography{refs}

>>>>

\documentclass[review,onefignum,onetabnum]{siamart190516}
%\usepackage{natbib}
\pdfoutput=1
%\usepackage[colorlinks=true,urlcolor=blue,citecolor=blue,linkcolor=blue]{hyperref}
\usepackage[english]{babel}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{amssymb}
\usepackage{tabularx}
\usepackage{quoting}
\usepackage{upquote}
\usepackage{subcaption}
\usepackage{multicol}
\usepackage[framemethod=TikZ]{mdframed}
\usetikzlibrary{shapes}
\usepackage{wrapfig}
%\usepackage{caption}
%\usepackage[plain]{algorithmic}
\usepackage[linesnumbered, ruled, vlined, algo2e]{algorithm2e}
\usepackage{algpseudocode}
\usepackage{rotating}
%\usepackage{cite}
\usepackage{booktabs}
%\usepackage{unicode-math}
%\usepackage{algorithm}% http://ctan.org/pkg/algorithm
%\usepackage{algpseudocode}% http://ctan.org/pkg/algpseudocode
\usepackage{xcolor}% http://ctan.org/pkg/xcolor
\makeatletter
\newsavebox{\@brx}
\newcommand{\llangle}[1][]{\savebox{\@brx}{\(\m@th{#1\langle}\)}%
  \mathopen{\copy\@brx\kern-0.5\wd\@brx\usebox{\@brx}}}
\newcommand{\rrangle}[1][]{\savebox{\@brx}{\(\m@th{#1\rangle}\)}%
  \mathclose{\copy\@brx\kern-0.5\wd\@brx\usebox{\@brx}}}
\makeatother
\usepackage{bbm}
\usepackage{jlcode}
\usepackage{graphicx}
\usepackage{amsmath,color}
\usepackage{mathrsfs}
\usepackage{float}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{indentfirst}
\usepackage{txfonts}
\usepackage[epsilon, tsrm, altpo]{backnaur}

@@

%%%%%% Shortcut related
\newcommand{\<}{\langle}
\renewcommand{\>}{\rangle}
\newcommand{\out}{{\vx^L}}
\newcommand{\inp}{{\vx^0}}
\newcommand{\cquad}{{{ }_{\quad}}}
\newcommand{\pluseq}{\mathrel{+}=}
\newcommand{\minuseq}{\mathrel{-}=}
\newcommand{\vx}{{\mathbf{x}}}
\newcommand{\vg}{{\mathbf{g}}}
\newcommand{\vp}{{\mathbf{p}}}
\newcommand{\vy}{{\mathbf{y}}}
\newcommand{\Var}{{\mathrm{Var}}}
\newcommand{\Mean}{{\mathrm{E}}}
\newcommand{\vvalue}{{\texttt{value}}}
\newcommand{\grad}{{\texttt{grad}}}
\newcommand{\parameter}{{\texttt{parameter}}}
%%%%%% Convention related
\newcommand{\SWAP}{{\rm SWAP}}
\newcommand{\CNOT}{{\rm CNOT}}
\newcommand{\bigO}{{\mathcal{O}}}
\newcommand{\X}{{\rm X}}
\renewcommand{\H}{{\rm H}}
\newcommand{\Rx}{{\rm Rx}}
\renewcommand{\v}[1]{{\bf #1}}
\newcommand{\dataset}{{\mathcal{D}}}
\newcommand{\wfunc}{{\psi}}
\newcommand{\SU}{{\rm SU}}
\newcommand{\UU}{{\rm U}}
\newcommand{\thetav}{{\boldsymbol{\theta}}}
\newcommand{\gammav}{{\boldsymbol{\gamma}}}
\newcommand{\thetai}{{\theta^\alpha_l}}
\newcommand{\Expect}{{\mathbb{E}}}
\newcommand{\Tr}{{\rm Tr}}
\newcommand{\etc}{{\it etc~}}
\newcommand{\etal}{{\it etal~}}
\newcommand{\xset}{\mathbf{X}}
\newcommand{\fl}{\texttt{fl}}
\newcommand{\pdata}{\mathbf{\pi}}
\newcommand{\q}{\mathbf{q}}
\newcommand{\epdata}{\mathbf{\hat{\pi}}}
\newcommand{\gammaset}{\boldsymbol{\Gamma}}
\newcommand{\ei}{{\mathbf{e}_l^\alpha}}
\newcommand{\vtheta}{{\boldsymbol{\theta}}}
\newcommand{\sigmag}{{\nu}}
\newcommand{\sigmai}[2]{{\sigma^{#2}_{#1}}}
\newcommand{\qi}[1]{{q^{\alpha_{#1}}_{#1}}}
\newcommand{\BAS}{Bars-and-Stripes}
\newcommand{\circled}[1]{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {#1}}}}
\newcommand{\qexpect}[1]{{\left\langle #1\right\rangle}}
\newcommand{\expect}[2]{{\mathop{\mathbb{E}}\limits_{\substack{#2}}\left[#1\right]}}
\newcommand{\var}[2]{{\mathop{\mathrm{Var}}\limits_{\substack{#2}}\left(#1\right)}}
\newcommand{\pshift}[1]{{p_{\thetav+#1}}}
\newcommand{\upcite}[1]{\textsuperscript{\cite{#1}}}
\newcommand{\Eq}[1]{Eq.~(\ref{#1})}
\newcommand{\Fig}[1]{Fig.~\ref{#1}}
\newcommand{\Lst}[1]{Listing.~\ref{#1}}
\newcommand{\Tbl}[1]{Table~\ref{#1}}
\newcommand{\Sec}[1]{Sec.~\ref{#1}}
\newcommand{\App}[1]{Appendix~\ref{#1}}
\newcommand{\bra}[1]{\mbox{$\left\langle #1 \right|$}}
\newcommand{\ket}[1]{\mbox{$\left| #1 \right\rangle$}}
\newcommand{\braket}[2]{\mbox{$\left\langle #1 | #2 \right\rangle$}}
\newcommand{\tr}[1]{\mathrm{tr}\mbox{$\left[ #1\right]$}}

@@

%\newtheorem{theorem}{\textit{Theorem}}
%\newtheorem{corollary}{\textit Branching Rule}
%\theoremstyle{definition}\newtheorem{definition}{\textit{Definition}}
%\newtheorem{defin}[thm]{Definition}

@@

%\externaldocument{ex_supplement}

\title{Solving the maximum independant set problem by generic programming einsum networks
\thanks{\funding{...}}
}

\author{XXX\thanks{XXX 
  (\email{email}, \url{website}).}
\and YYY\thanks{yyyyy 
  (\email{yyyy}, \email{email}).}
}

\begin{document}

\maketitle

\begin{abstract}
	Solving the maximum independent set size problem by mapping the graph to an einsum network. 
    We show how to obtain the maximum independent set size, the independence polynomial and optimal configurations of a graph by engineering the tensor element algebra.
\end{abstract}

\section{Introduction}
MIS problem is hard~\cite{Hastad1996}.
Branching algorithms can solve the MIS problem in $1.1893^n n ^{O(1)}$ time ~\cite{Xiao2017}.
Previous dynamic programming approach~\cite{Fomin2013} can reduce the complexity of computing to $2^{tw(G)}$.
A set is independent if and only if it is a clique in the graph's complement, so the two concepts are complementary.
A set is independent if and only if its complement is a vertex cover. Therefore, the sum of the size of the largest independent set $\alpha (G)$ and the size of a minimum vertex cover $\beta (G)$ is equal to the number of vertices in the graph.
It is related to hard spheres lattice gas model~\cite{Dyre2016}, and Rydberg hamiltonian~\cite{Pichler2018}.

In this work, we attack this problem by mapping it to an ``einsum'' network.
The word ``einsum'' is a shorthand for Einstein's summation, however, modern einsum notation in program is actually invented by a group of programmers.
Einstein's notation is originally proposed as a generalization to of multiplication between two matrices to the contraction between multiple tensors.
Let $A, B$ be two matrices, the matrix multiplication is defined as $C_{ik} = \sum_{j}A_{ij}B_{jk}$.
It is denoted as $C_i^k = A_i^j B_j^k$ in the Einstein's original notation, where the paired subscript and superscript $j$ is a dummy index summed over.
An example of tensor networks is $C_i^l = A_{ij}^k B^l_k V^j$.
One can map a tensor network to a muti-graph with open edges by viewing a tensor in the expression on the right hand side as a vertex in a graph, a label pairing two tensors as an edge, and the remaining labels as open edges.
We get the graphical notation as the following.

@@

\begin{algorithm}[!ht]
    \small
    \SetAlgoNoLine
    \LinesNumbered
    Let $P = 1$, vector $X = (0,1,2,\ldots,m)$, matrix $\hat X_{ij} = X_i^j$, where $i,j = 0, 1, \ldots m$\;

    \While{true}{
        compute the largest prime $p$ that $\gcd(p, P) = 1 \land p \leq 2^W$\;

        compute the tensor network contraction on $GF(p)$ and obtain $Y = (y_0, y_1, \ldots , y_m) \pmod p $\;

        $A_p = (a_0, a_1, \ldots, a_m) \pmod p = {\rm gaussian\_elimination}(\hat X, Y \pmod p) $\;

        $A_{P\times p} = {\rm chinese\_remainder}(A_P, A_p)$\;

        \If{$A_P = A_{P \times p}$}{
            \Return $A_P$ \tcp*[l]{converged}
        }
        $P = P \times p$\;
    }\caption{Compute independence polynomial exactly without integer overflow}\label{alg:finitefield}
\end{algorithm}

@@

\begin{theorem}[]
    A MIS-compact tropical tensor is optimal, i.e. none of its none zero entries can be removed without accessing global information.
\end{theorem}

@@

\bibliographystyle{siamplain}
\bibliography{refs}
